### 2015-08-19 paper-1/main.tex
Working through comments.

Our goal is to know the distribution of the sample size of the design for 
planning purposes.  If all 12 patients are enrolled at once, as in the classic 
design, then the sample size is 12. However, in most clinical trials, the 
patients are enrolled sequentially, often with one patient's outcome realized 
before the next one enters the trial. In the present example, observing two 
successful patients allows us to proceed to the next stage. Similarly 11 
observed treatment failures also ends the stage. This sampling mechanism, in 
which the experiment ends as soon as any of the endpoints is reached, is 
call {\em curtailed sampling}. Under curtailed sampling the range of the 
sample size is between two and 12.

@@

The next section of this paper introduces our notation and basic results including the density of the SNB and compound SNB along with a description of it's relation to other distributions. Sections 2 and 3 derive the distribution based on a defined Bernoulli process and give some basic properties of the distribution. Section 4 derives the compound distribution using a Beta prior. Section 5 develops a post-hoc analysis of a completed trial. Section 6 is devoted to discussion and conclusions. The simulations, model fitting routines, and visualizations presented in this paper were generated using the snb package [\ref{Kane 2015}] for the R programming environment [\ref{R Core Team 2015}]

@@

\begin{prop}
The compound PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
\begin{align}
f(k | s, t, \alpha, \beta) &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function.
\end{prop}
\begin{proof}
For notational simplicity, assume that $s,t \leq k \leq s+k-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

@@

Like most Bayesian Adaptive Clinical trials the SNB approach allows for monitoring of trial and can be updated as new results are received without losing power. Essentially this is because the object under investigation is not a point estimate of the success probability. It is the uncertainty in the distribution of the success probability. Point estimates of the success probability can be regarded as artifacts of our parameterization of the underlying distribution. Every point estimate calculated is accompanied by the quantification of the uncertainty around it. This estimate and it's uncertainty can be calculated at any time and may be useful in evaluating ongoing trials as well as providing a post-hoc evaluation of trials that have been completed.

@@

Fig. \ref{fig:conditional_snb} shows the compound SNB distribution. In each of the following enrollment a success would end the trial. If none of the following enrollments are success then the trial will stop at the eleventh patient. From Fig. \ref{fig:hypo} we know that the estimate of $p$ is about 0.19. The compound distribution gives the probability of another success as 0.3883 and the probability of no more success (and the first stage failing) as 0.6117. Furthermore, we can see that the probability that the trial ends in the next step is approximately 20\% and the probability the trials end with the $11^{th}$ patient is about 65\%.

@@

Suppose that, in our example trial, a second successful outcome was reached 
on the $10^{th}$ patient. From the data collected we have shown how to create 
an estimate for the distribution of $p$. Based on that estimate we may like to 
see the probability of a different outcome. We may want to answer the
question, "What is the probability that we don't get two successful outcomes?" 
We may also ask what is the distribution for the number of patients 
enrolled for a similar study?

>>>>

Our goal is to know the distribution of the sample size of the design for 
planning purposes and to characterize the uncertainty in the estimate of the
success probability.  If all 12 patients are enrolled at once, 
as in the classic 
design, then the sample size is 12. However, in most clinical trials, the 
patients are enrolled sequentially, often with one patient's outcome realized 
before the next one enters the trial. In the present example, observing two 
successful patients allows us to proceed to the next stage. Similarly 11 
observed treatment failures also ends the stage. This sampling mechanism, in 
which the experiment ends as soon as any of the endpoints is reached, is 
call {\em curtailed sampling}. Under curtailed sampling the range of the 
sample size is between two and 12.

@@

The next section of this paper introduces our notation and basic results 
including the density of the distribution along with a description of 
it's relation to other distributions. Sections 2 and 3 derive the distribution 
based on a defined Bernoulli process and give some basic properties of the 
distribution. Section 4 derives the compound distribution using a Beta prior. 
Section 5 develops a post-hoc analysis of a completed trial. Section 6 is 
devoted to discussion and conclusions. The simulations, model fitting 
routines, and visualizations presented in this paper were generated using the 
snb package [\ref{Kane 2015}] for the R programming 
environment [\ref{R Core Team 2015}]

@@

\begin{prop}
The compound PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
\begin{align}
f(k | s, t, \alpha, \beta) &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function.
\end{prop}
\begin{proof}
For notational simplicity, assume that $s,t \leq k \leq s+k-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

@@

Like most Bayesian Adaptive Clinical trials the SNB approach allows for monitoring of trial and can be updated as new results are received without losing 
power. This is because the object under investigation is not a point estimate 
of the success probability; it is the uncertainty in the distribution of the 
success probability. Point estimates of the success probability can be 
regarded as artifacts of our parameterization of the underlying distribution. 
Every point estimate calculated is accompanied by the quantification of the 
uncertainty around it. This estimate and its uncertainty can be calculated at 
any time and may be useful in evaluating ongoing trials as well as providing a 
post-hoc evaluation of trials that have been completed.

@@

Fig \ref{fig:hypo} shows that at the eight enrollment the estimate of $p$ is 
about 0.19 and, from the lower 5\% quantile, we can see that there is
a 95\% probability that $p$ is greater than 2.5\%. 
Fig. \ref{fig:conditional_snb} shows the compound SNB distribution for this 
hypothetical trial at the eighth step.  Successes may occur 
at steps nine through 11 with probability 26.7\%. The remaining probability
mass corresponds to reaching the $11^{th}$ step without a second success.

@@

Suppose that, in our example trial, a second successful outcome was reached 
on the $10^{th}$ patient. From the data collected we have shown how to create 
an estimate for the distribution of $p$. Based on that estimate we may like to 
see the probability of a different outcome. We may want to answer the
question, ``What is the probability that we don't get two successful outcomes?''
We may also ask what is the distribution for the number of patients 
enrolled for a similar study?
